Transition path sampling based calculations of free energies of enzymatic reactions: the case of human methionine adenosyl transferase and plasmodium vivax adenosine deaminase

### 2022-05-05 tex/manuscript.tex
added some more text

The reaction catalyzed by the human MAT2A enzyme is
between adenosyl $5'$-triphosphate (ATP) and L-methionine (MET) molecules resulting in the formation of
S-adenosyl-L-methionine (SAM) which is depicted in Fig. \ref{fig:mat2a-reaction}.
 This reaction was experimentally characterized by 
Firestone et al. \cite{Firestone17JAmChemSoc139p13754} to follow a S$_{\text{N}}2$ transition 
state. The key bond parameters for this reaction consists of the distance between the
nucleophilic sulfur atom of MET and the $5'-C$ atom of the ATP (denoted as $dis(SC)$), and the distance between 
oxygen atom of the phosphate group and the $5'-C$ atom of ATP ($dis(OC)$).
For the TPS calculations we define the reactant state to have $dis(OC)$ to be 
less than 1.7 {\AA} and the product state to have the $dis(SC)$ less than 1.7 {\AA}.

@@

\begin{figure}
\includegraphics[scale=1.0]{figures/mat2a-diff167.pdf}
\caption{Bond breaking ($dis(OC)$), bond forming ($dis(SC)$) distances for a typical 
reactive trajectory for the reaction catalyzed by the MAT2A enzyme.}
\label{fig:mat2a-reactive-traj}
\end{figure}
The order parameters of the system along a typical reactive trajectory is 
shown in Fig. \ref{fig:mat2a-reactive-traj}.

@@

%\begin{figure}[ht!]
%\includegraphics[scale=0.2]{figures/mat2a-transition-state.png}
%\includegraphics[scale=0.2]{figures/mat2a-active-site2.png}
%\end{figure}

@@

\caption{A reactive trajectory}

@@

The MAT2A enzymatic reaction was analyzed in detail to elucidate the transition state,
reaction coordinate and the free energy profile. The reaction proceeds through a 
S$_\text{N}$2 mechanism, the d$_{\text{OC}}$ and d$_{\text{SC}}$ are $2.1$ {\AA}
and $2.3$ {\AA}, respectively. The reaction coordinate is determined to involve  
the Gln113, Ser114, Arg249 and Arg264 residues. 
The free energy barrier for this reaction was calculated to be 
$16.0\;\text{Kcal}\;\text{mol}^{-1}$ which is in good agreement 
with the experimentally measured activation energy of $17.27\;\pm\;1.5\;\text{Kcal mol}^{-1}$. 
Furthermore the reaction catalyzed by the $pv$ADA enzyme is analyzed in detail to obtain the  
transition state and calculate the free energy profile.
The transition state has a Meisenheimer complex like structure with the 
d$_\text{NC}$ and d$_\text{OC}$ being $1.60$ {\AA} and $1.48$ {\AA}, respectively.
The free energy barrier for this reaction for found to be $23\;\text{Kcal}\;\text{mol}^{-1}$
with a double well shape similar to that measured from experiments.  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The "Acknowledgement" section can be given in all manuscript
%% classes.  This should be given within the "acknowledgement"
%% environment, which will make the correct section or running title.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{acknowledgement}

>>>>

The reaction catalyzed by the human MAT2A enzyme is
between adenosyl $5'$-triphosphate (ATP) and L-methionine (MET) molecules 
resulting in the formation of S-adenosyl-L-methionine (SAM) which is 
depicted in Fig. \ref{fig:mat2a-reaction}. This reaction follows a S$_{\text{N}}2$ 
mechanism and was experimentally characterized by 
Firestone et al. \cite{Firestone17JAmChemSoc139p13754}. The key bond 
parameters for this reaction consists of the distance between the
nucleophilic sulfur atom of MET and the $5'-C$ atom of the ATP (denoted as d$_{\mathrm{SC}}$), 
and the distance between oxygen atom of the phosphate group and the $5'-C$ 
atom of ATP (d$_{\mathrm{OC}}$). For the TPS calculations we define the 
reactant state to have d$_{\mathrm{OC}}$ to be less than 1.7 {\AA} and the product 
state to have the d$_{\mathrm{SC}}$ less than 1.7 {\AA}.

@@

\begin{figure}
\includegraphics[scale=1.0]{figures/mat2a-diff167.pdf}
\caption{Bond breaking ($\mathrm{d}_{\mathrm{OC}}$), bond forming 
($\mathrm{d}_{\mathrm{SC}}$) distances for a typical 
reactive trajectory for the reaction catalyzed by the MAT2A enzyme.}
\label{fig:mat2a-reactive-traj}
\end{figure}
The order parameters of the system along a typical reactive trajectory is 
shown in Fig. \ref{fig:mat2a-reactive-traj}.

@@

@@

\caption{Bond breaking ($\mathrm{d}_{\mathrm{NC}}$), bond forming 
($\mathrm{d}_{\mathrm{OC}}$) distances for a typical 
reactive trajectory for the reaction catalyzed by the $pv$ADA enzyme.}

@@

The MAT2A enzymatic reaction was analyzed in detail to elucidate the transition state,
reaction coordinate and the free energy profile. The reaction proceeds through a 
S$_\text{N}$2 mechanism, the d$_{\text{OC}}$ and d$_{\text{SC}}$ are $2.1$ {\AA}
and $2.3$ {\AA}, respectively. The reaction coordinate is determined to involve  
the Gln113, Ser114, Arg249 and Arg264 residues. 
The free energy barrier for this reaction was calculated to be 
$16.0\;\text{Kcal}\;\text{mol}^{-1}$ which is in good agreement 
with the experimentally measured activation energy of $17.27\;\pm\;1.5\;\text{Kcal mol}^{-1}$. 
Furthermore the reaction catalyzed by the $pv$ADA enzyme is analyzed in detail to obtain the  
transition state and calculate the free energy profile.
The transition state has a Meisenheimer complex like structure with the 
d$_\text{NC}$ and d$_\text{OC}$ being $1.60$ {\AA} and $1.48$ {\AA}, respectively.
The free energy barrier for this reaction for found to be $23\;\text{Kcal}\;\text{mol}^{-1}$
with a double well shape similar to that measured from experiments.  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The "Acknowledgement" section can be given in all manuscript
%% classes.  This should be given within the "acknowledgement"
%% environment, which will make the correct section or running title.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{acknowledgement}
SGB acknowledges funding from ...
